Dynamical Properties of Molecular Cloud Complexes in a Redshift 6 Prototypical Galaxy at the Epoch of Reionization

### 2018-09-12 WRITEUP/ms.tex
clean up and hide toomre Q section for report

This paper is structured as follows.
% \AP{TBM}
In \Sec{sim}, we describe the setup of our simulation and properties of our main galaxy (\flower),
and describe the method used to identify its molecular gas complexes.
In \Sec{results}, we present scaling relations based on
the MC identified and compare them with observations of molecular clouds
and structures seen in nearby and \z$>$\,0 galaxies.
We present Toomre stability analysis in \Sec{Q}.
We discuss the results % , robustness of our analysis, caveats, 
and implications of our findings in \Sec{diss},
and present our conclusions in \Sec{conclusion}.
Throughout this paper, we adopt a concordance cosmology, with total matter, vacuum and baryonic densities
in units of the critical density $\Omega_{\Lambda}$\eq0.692, $\Omega_m$\eq0.308, $\Omega_b$\eq0.0481,
Hubble constant $H_0$\eq100\,$h$\,km s\pmOne\,Mpc\pmOne with $h$\eq0.678,
spectral index $n$\eq0.967 and $\sigma_8$\eq0.826 \citep{Planck14a}.

@@

\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em] 
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples showing the MCs identified in our simulation by applying volumetric H$_2$ density cuts of
$n_{\rm cut}$\eq[0.53, 6.81,19.00]\,cm$^{-3}$, which is only a subset of all the $n_{\rm cut}$ adopted (see text).
	Color shows the H$_2$ surface density, weighted by the column density. Specifically, this is calculated using \Eq{integrate}.
% 	\AP{plot a proper label for the colorbar, and remove the trailing right axes labels (that are still there despite our efforts)}\DL{I tried more on Sept 7th, errrr, now all the ticks and labels for both panels disappear... hmm.. Not worth my time to fix this for the report.}
\label{fig:MC}}
% \addtocounter{figure}{-1}
\end{figure*}

@@

For visualization, we overplot the molecular structures identified using
a subset of the $n_{\rm cut}$ (0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps 
in \Fig{MC} as an example.
The plots shown in \Fig{MC} are weighted by the total gas density; that is, 
\begin{equation}
{\rm pixel_i} = \frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}, 
\label{eqn:integrate}
\end{equation} 
where $\rho_{\rm H2, i}$ is the H$_2$ density of pixel $i$, $w_i$ is the weighted quantity in pixel $i$ which we adopt 
to be the total gas density, and $\Delta z_i$ is the path length of pixel $i$ along the line-of-sight over which the quantities are integrated along.

@@

The mass of an MC is calculated from the uniformly-gridded 3D density field,
integrating over all zones of the MC, multiplied by its volume. 
The size of an MC ($R$) is defined assuming spherical geometry (i.e., 
the size is parameterized via the radius of a sphere, which has a volume corresponding to that of the identified MC).
% {\AP The size is an effective radius, calculated from the total volume of the MC assuming spherical symmetry.}
Since in \obs, dense gas contributes more to the observed linewidths than the more diffuse gas. We therefore calculate 
the non-thermal component of the velocity dispersion ($\sigma_{\rm NT}$) of each MC as a density-weighted quantity:
\begin{equation}
\sigma_{\rm NT}^2 = \frac{1}{{3}}\frac{\sum_{i} \rho_i \left|\mathbf{v}_i - \mathbf{\bar v}\right|^2}{\sum_i \rho_i}\,,
\end{equation}
which is summed over the cells composing each MC.
The total velocity dispersion includes contributions from the thermal sound speed ($c_s$):
\begin{equation}
\sigma^2 = \sigma_{\rm NT}^2 + c_s^2.
\end{equation}
The local sound speed ($c_s$)
of all identified molecular complexes are generally much smaller compared to the their
turbulent velocities\footnote{We find comparable velocity dispersion between that derived from 
taking the root mean square versus that derived from thermal and non-thermal pressure terms, indicating that
rotation velocity is unlikely to be the dominant source of velocity dispersion reported here.
We further quantify the contributions arising from galactic rotation and shear to the extracted velocity dispersions of the MCs in \Sec{Q}.
In the subsequent sections, we adopt the root mean square of the velocity field
as the velocity dispersion ($\sigma$).}.
On average, the Mach number of the MCs corresponds to about $\mathcal{M}\approx50$.
This is consistent with the global analysis done on \flower by \citet{Vallini18a}.

@@

The lower $\alpha_{\rm vir}$ and $P_{\rm ext}$ compared to their parent MCs are consistent with
the conventional wisdom that collapsing structures result from local gravitational
instability within globally non-collapsing structures \citep[see e.g.,][]{Ballesteros-Paredes11a},
which are supported by turbulence and rotation. 
In fact, based on the virial parameter, velocity dispersion, and pressure of the MC identified in the main disk of \flower, 
the MC itself is gravitationally unbound and is supported
by turbulence (from the feedback of multiple episodes of \SF) 
and galactic rotation on large scales (occupying the high pressure region shown in the bottom panel of \Fig{alpha16}).
% Turbulence in these substructures further dissipates in % e.g., shocks on even smaller scales, allowing sub-MCs to be bound and collapse.

@@

\section{Origin of the clumps: Mass Scales set by Toomre and Jeans Instabilities}   \label{sec:Q}
The onset of gravitational instability is considered to be tightly connected to (the conditions for) \SF \citep[e.g.,][]{Kennicutt89a, Li05b, Li06a}
and is widely discussed in the context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}.
For axisymmetric mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk
of finite thickness $h$
is described by:
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber in the disk and $\kappa$ is the epicyclic frequency.
The first term on the RHS describes rotation support, second term describes self-gravity, and
the third term describes pressure support.
The epicyclic frequency is defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right).
\label{eqn:kappa}
\end{equation}
% \kappa\propto\Omega
One can then define the so-called Toomre-Q parameter/criterion for stability as
\begin{equation}
Q\equiv\frac{\sigma\kappa}{\pi G \Sigma}.
\label{eqn:Q}
\end{equation}
We can then rewrite Equation~\ref{eqn:3Ddisp} in terms of
\begin{equation}
\omega^2 = Q^2 k^2 \left(\frac{\pi G \Sigma}{\kappa}\right)^2 - 2\pi G \Sigma |k| + \kappa^2.
\end{equation}
Thus, for $Q\lesssim$1, disk is unstable for some $k$ ($\omega^2 < 0$)
and fragmentation can happen
(e.g., instability impose by angular momentum from gas accretion from the cosmic web and
satellite galaxies and stellar feedback).
That is, local fragmentation takes place when gravity overcomes pressure support
from turbulence (thermal is small here) and shearing of differential rotation.

In the limit of a thin rotating disk, rotation can help stabilizes
self-gravitational contraction for wavelengths greater than the Toomre length:
\begin{equation}
\lambda_T = 4\pi^2 G\Sigma/\kappa^2,
\end{equation}
yielding a characteristic mass of $M_T$\eq$\frac{\pi}{4}{\lambda_T^2}{\Sigma}$\eq4$\pi^2\frac{G^2\Sigma^3}{\kappa^4}$,
which sets the ``upper limit'' to which MCs can form in a shearing disk.
\DL{still yet to make the ``Q''-map}

We overplot the MCs on the Toomre-Q map.

>>>>

This paper is structured as follows.
% \AP{TBM}
In \Sec{sim}, we describe the setup of our simulation and properties of our main galaxy (\flower),
and describe the method used to identify its molecular gas complexes.
In \Sec{results}, we present scaling relations based on
the MC identified and compare them with observations of molecular clouds
and structures seen in nearby and \z$>$\,0 galaxies.
% We present Toomre stability analysis in \Sec{Q}.
We discuss the results % , robustness of our analysis, caveats, 
and implications of our findings in \Sec{diss},
and present our conclusions in \Sec{conclusion}.
Throughout this paper, we adopt a concordance cosmology, with total matter, vacuum and baryonic densities
in units of the critical density $\Omega_{\Lambda}$\eq0.692, $\Omega_m$\eq0.308, $\Omega_b$\eq0.0481,
Hubble constant $H_0$\eq100\,$h$\,km s\pmOne\,Mpc\pmOne with $h$\eq0.678,
spectral index $n$\eq0.967 and $\sigma_8$\eq0.826 \citep{Planck14a}.

@@

\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em] 
  \includegraphics[scale=0.65]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples showing the MCs identified in our simulation by applying volumetric H$_2$ density cuts of
$n_{\rm cut}$\eq[0.53, 6.81,19.00]\,cm$^{-3}$, which is only a subset of all the $n_{\rm cut}$ adopted (see text).
	Color shows the H$_2$ surface density, weighted by the column density. 
% 	\AP{plot a proper label for the colorbar, and remove the trailing right axes labels (that are still there despite our efforts)}\DL{I tried more on Sept 7th, errrr, now all the ticks and labels for both panels disappear... hmm.. Not worth my time to fix this for the report.}
\label{fig:MC}}
% \addtocounter{figure}{-1}
\end{figure*}

@@

For visualization, we overplot the molecular structures identified using
a subset of the $n_{\rm cut}$ (0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps 
in \Fig{MC} as an example.
The plots shown in \Fig{MC} are weighted by the column density of the total gas content.
% equation below for bookkeeping: 
% ${\rm pixel_i}$\eq$\frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}$,
% where $\rho_{\rm H2, i}$ is the H$_2$ density of pixel $i$, $w_i$ is the weighted quantity in pixel $i$ which we adopt 
% to be the total gas density, and $\Delta z_i$ is the path length of pixel $i$ along the line-of-sight over which the quantities are integrated along.

@@

The mass of an MC is calculated from the uniformly-gridded 3D density field,
integrating over all zones of the MC, multiplied by its volume. 
The size of an MC ($R$) is defined assuming spherical geometry (i.e., 
the size is parameterized via the radius of a sphere, which has a volume corresponding to that of the identified MC).
% {\AP The size is an effective radius, calculated from the total volume of the MC assuming spherical symmetry.}
Since in \obs, dense gas contributes more to the observed linewidths than the more diffuse gas. We therefore calculate 
the non-thermal component of the velocity dispersion ($\sigma_{\rm NT}$) of each MC as a density-weighted quantity:
\begin{equation}
\sigma_{\rm NT}^2 = \frac{1}{{3}}\frac{\sum_{i} \rho_i \left|\mathbf{v}_i - \mathbf{\bar v}\right|^2}{\sum_i \rho_i}\,,
\end{equation}
which is summed over the cells composing each MC.
The total velocity dispersion includes contributions from the thermal sound speed ($c_s$):
\begin{equation}
\sigma^2 = \sigma_{\rm NT}^2 + c_s^2.
\end{equation}
The local sound speed ($c_s$)
of all identified molecular complexes are generally much smaller compared to the their
turbulent velocities\footnote{We find comparable velocity dispersion between that derived from 
taking the root mean square versus that derived from thermal and non-thermal pressure terms, indicating that
rotation velocity is unlikely to be the dominant source of velocity dispersion reported here.
% We further quantify the contributions arising from galactic rotation and shear to the extracted velocity dispersions of the MCs in \Sec{Q}.
In the subsequent sections, we adopt the root mean square of the velocity field
as the velocity dispersion ($\sigma$).}.
On average, the Mach number of the MCs corresponds to about $\mathcal{M}\approx50$.
This is consistent with the global analysis done on \flower by \citet{Vallini18a}.

@@

The lower $\alpha_{\rm vir}$ and $P_{\rm ext}$ compared to their parent MCs are consistent with
the notion that collapsing structures result from local gravitational
instability within globally non-collapsing structures \citep[see e.g.,][]{Ballesteros-Paredes11a},
which are supported by turbulence and rotation. 
In fact, based on the virial parameter, velocity dispersion, and pressure of the MC identified in the main disk of \flower, 
the MC itself is gravitationally unbound and is supported
by turbulence (from the feedback of multiple episodes of \SF) 
and galactic rotation on large scales (occupying the high pressure region shown in the bottom panel of \Fig{alpha16}).
% Turbulence in these substructures further dissipates in % e.g., shocks on even smaller scales, allowing sub-MCs to be bound and collapse.

@@

%
%\section{Origin of the clumps: Mass Scales set by Toomre and Jeans Instabilities}   \label{sec:Q}
%The onset of gravitational instability is considered to be tightly connected to (the conditions for) \SF \citep[e.g.,][]{Kennicutt89a, Li05b, Li06a}
%and is widely discussed in the context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}.
%For axisymmetric mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk
%of finite thickness $h$
%is described by:
%\begin{equation}
%\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2,
%\label{eqn:3Ddisp}
%\end{equation}
%where $k$ is the wavenumber in the disk and $\kappa$ is the epicyclic frequency.
%The first term on the RHS describes rotation support, second term describes self-gravity, and
%the third term describes pressure support.
%The epicyclic frequency is defined as:
%\begin{equation}
%\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right).
%\label{eqn:kappa}
%\end{equation}
%% \kappa\propto\Omega
%One can then define the so-called Toomre-Q parameter/criterion for stability as
%\begin{equation}
%Q\equiv\frac{\sigma\kappa}{\pi G \Sigma}.
%\label{eqn:Q}
%\end{equation}
%We can then rewrite Equation~\ref{eqn:3Ddisp} in terms of
%\begin{equation}
%\omega^2 = Q^2 k^2 \left(\frac{\pi G \Sigma}{\kappa}\right)^2 - 2\pi G \Sigma |k| + \kappa^2.
%\end{equation}
%Thus, for $Q\lesssim$1, disk is unstable for some $k$ ($\omega^2 < 0$)
%and fragmentation can happen
%(e.g., instability impose by angular momentum from gas accretion from the cosmic web and
%satellite galaxies and stellar feedback).
%That is, local fragmentation takes place when gravity overcomes pressure support
%from turbulence (thermal is small here) and shearing of differential rotation.
%
%In the limit of a thin rotating disk, rotation can help stabilizes
%self-gravitational contraction for wavelengths greater than the Toomre length:
%\begin{equation}
%\lambda_T = 4\pi^2 G\Sigma/\kappa^2,
%\end{equation}
%yielding a characteristic mass of $M_T$\eq$\frac{\pi}{4}{\lambda_T^2}{\Sigma}$\eq4$\pi^2\frac{G^2\Sigma^3}{\kappa^4}$,
%which sets the ``upper limit'' to which MCs can form in a shearing disk.
%\DL{still yet to make the ``Q''-map}
%
%We overplot the MCs on the Toomre-Q map.
